---
title: "Steph Cury and NBA 3 Point Analysis"
author: "Jacob Weiss and Seigo Tomozawa"
date: '2022-06-08'
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(ggplot2)
library(lubridate)
library(rvest)
library(janitor)
library(plotly)
library(stringr)
library(scales)
library(xml2)
library(hoopR)
library(nbastatR)
library(purrr)
library(sportyR)
theme_set(theme_bw())
```

## Introduction

-   Since Stephen Curry has been in the NBA, 3-point shooting has increased substantially while midrange shots have steadily decreased
-   We decided to look at how Stephen Curry compares to other 3 point shooters
-   We also looked at in the past 23 years how 3 point shooting has changed

## **Stephen Curry compared to All Time Top 20 NBA 3 Point Shooters**

-   We first decided to look at the top 20 NBA 3 point shooters all time and compare them to Stephen Curry

-   This scatter graph shows the each of the top 20 three point shooters by the amount of three pointers made and the number of games played

    ```{r, include=FALSE}
    reg_url2 <-"https://www.landofbasketball.com/all_time_leaders/3_pointers_total_career_season.htm"
    pg3 <- xml2::read_html(reg_url2)
      basketball_three_leaders <- rvest::html_table(pg3, fill = TRUE)[[1]]
      names(basketball_three_leaders) <- basketball_three_leaders[1, ]
    basketball_three_leaders <- basketball_three_leaders[c(2:11), ]
      basketball_three_leaders_2 <- janitor::clean_names(basketball_three_leaders)

    basketball_three_leaders_2 <- basketball_three_leaders_2 %>%
      select(., player, x3_pointers_made, seasons, games, total3_pointers_attempted)
    basketball_three_leaders_2$x3_pointers_made <- as.numeric(gsub(",","",basketball_three_leaders_2$x3_pointers_made))
    basketball_three_leaders_2$games <- as.numeric(gsub(",","",basketball_three_leaders_2$games))
    basketball_three_leaders_2$total3_pointers_attempted <- as.numeric(gsub(",","",basketball_three_leaders_2$total3_pointers_attempted))
    basketball_three_leaders_2[, c(2, 3, 4, 5)] <- sapply(basketball_three_leaders_2[ c(2, 3, 4, 5)], as.numeric )

    pg3 <- xml2::read_html(reg_url2)
      basketball_three_leaders_3 <- rvest::html_table(pg3, fill = TRUE)[[2]]
      names(basketball_three_leaders_3) <- basketball_three_leaders_3[1, ]
    basketball_three_leaders_3 <- basketball_three_leaders_3[c(2:11), ]
      basketball_three_leaders_4 <- janitor::clean_names(basketball_three_leaders_3)

    basketball_three_leaders_4 <- basketball_three_leaders_4 %>%
      select(., player, x3_pointers_made, seasons, games, total3_pointers_attempted)
    basketball_three_leaders_4$x3_pointers_made <- as.numeric(gsub(",","",basketball_three_leaders_4$x3_pointers_made))
    basketball_three_leaders_4$games <- as.numeric(gsub(",","",basketball_three_leaders_4$games))
    basketball_three_leaders_4$total3_pointers_attempted <- as.numeric(gsub(",","",basketball_three_leaders_4$total3_pointers_attempted))
    basketball_three_leaders_4[, c(2, 3, 4, 5)] <- sapply(basketball_three_leaders_4[ c(2, 3, 4, 5)], as.numeric )

    total_3 <- rbind(basketball_three_leaders_2, basketball_three_leaders_4)
    ```

```{r pressure}
p <- ggplot(data = total_3,
            aes(x = games, y = x3_pointers_made,col = player,
                text = paste('Players: ', player, '\n',
                             '3 Pointers Made:', x3_pointers_made, '\n', 'Games Played: ', games))) +
  geom_point() + 
  labs(title = "Top 20 Career 3 Pointers Made", x = "Games", y = "Three Pointers Made")
ggplotly(p, tooltip = "text")
```

## Stephen Curry Shooting Chart Compaired to Ray Allen

-   We then decided to take the analysis a step farther by comparing Stephen Curry to the second best three point shooter Ray Allen

-   We used the nbastatR package

-   The graphs for both Ray Allen and Stephen Curry are of their shooting positions when taking 3 point attempts

    ```{r, include=FALSE}
    Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2)
    warriors_shot_chart_df <- dplyr::bind_rows(
      nbastatR::teams_shots(teams = "Bucks", seasons = 2000),
      nbastatR::teams_shots(teams = "Bucks", seasons = 2001),
      nbastatR::teams_shots(teams = "Bucks", seasons = 2002),
      nbastatR::teams_shots(teams = "Celtics", seasons = 2007),
      nbastatR::teams_shots(teams = "Celtics", seasons = 2008),
      nbastatR::teams_shots(teams = "Celtics", seasons = 2009),
      nbastatR::teams_shots(teams = "Celtics", seasons = 2010),
      nbastatR::teams_shots(teams = "Celtics", seasons = 2011),
      nbastatR::teams_shots(teams = "Celtics", seasons = 2012),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2010),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2011),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2012),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2013),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2014),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2015),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2016),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2017),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2018),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2019),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2020),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2021),
      nbastatR::teams_shots(teams = "Warriors", seasons = 2022))
    ```

For Stephen Curry:

```{r curry}
p <- geom_basketball(league = "NBA", full_surf = F, rotate = T,
court_background_color = "white",
court_apron_color = "white",
two_point_range_color = "white",
amateur_painted_area_color = "white",
professional_painted_area_color = "white",
free_throw_semi_circle_fill_color = "white")
p + geom_point(data = warriors_shot_chart_df %>%
dplyr::filter(., zoneBasic != "Backcourt",
yearSeason %in% c(2010:2022),
namePlayer == "Stephen Curry",
typeShot == "3PT Field Goal"),
aes(x = -locationX/10, y = (locationY - 420)/10,
col = zoneBasic),
alpha = .15) +
scale_color_brewer("Zone", palette = "Dark2") +
labs(caption = "")
```

For Ray Allen:

```{r Allen}
p + geom_point(data = warriors_shot_chart_df %>%
dplyr::filter(., zoneBasic != "Backcourt",
yearSeason %in% c(2000:2012),
namePlayer == "Ray Allen",
typeShot == "3PT Field Goal"),
aes(x = -locationX/10, y = (locationY - 420)/10,
col = zoneBasic),
alpha = .15) +
scale_color_brewer("Zone", palette = "Dark2") +
labs(caption = "")
```

## NBA 3 Point Shooting

-   While thinking about how the NBA overall 3 point shooting has changed we decided to visualize this

-   We pulled our data from basketballrefrence.com

-   We used the regular season averages for each team to look at how the 3 point shooting rate has changed since 1999 to 2022

    ```{r, include=FALSE}
    results <- dplyr::tibble()
    #adv_url <- "https://www.basketball-reference.com/leagues/NBA_1999.html"
    years <- 1999:2015
    for(i in seq_along(years)){
      cat(sprintf('Year: %s at %s\n', years[i], Sys.time()))
      reg_url <- paste("https://www.basketball-reference.com/leagues/NBA_", years[i],
                       ".html", sep = "")
      pg <- xml2::read_html(reg_url)
      basketball_reg_stats <- rvest::html_table(pg, fill = TRUE)[[3]]
      basketball_reg_stats <- dplyr::mutate(basketball_reg_stats, season = years[i])
      results <- dplyr::bind_rows(results, basketball_reg_stats)
    }

    results1 <- dplyr::tibble()
    #adv_url <- "https://www.basketball-reference.com/leagues/NBA_1999.html"
    years1 <- 2016:2022
    for(k in seq_along(years1)){
      cat(sprintf('Year: %s at %s\n', years1[k], Sys.time()))
      reg_url1 <- paste("https://www.basketball-reference.com/leagues/NBA_", years1[k],
                       ".html", sep = "")
      pg <- xml2::read_html(reg_url1)
      basketball_reg_stats1 <- rvest::html_table(pg, fill = TRUE)[[5]]
      basketball_reg_stats1 <- dplyr::mutate(basketball_reg_stats1, season = years1[k])
      results1 <- dplyr::bind_rows(results1, basketball_reg_stats1)
    }
    results_final <- rbind(results, results1)


    saveRDS(results_final, "\\Users\\jacob\\Info 3700 (Sports Statistics)\\Project Info 3700\\nba-basketball-stats-1999-2021.rds")

    ```

    ```{r season}
    Basketball_Data <- readRDS("\\Users\\jacob\\Info 3700 (Sports Statistics)\\Project Info 3700\\nba-basketball-stats-1999-2021.rds")
    Basketball_Data_Filtered <- Basketball_Data %>%
      filter(., Team == 'League Average') 
    colnames(Basketball_Data_Filtered)[9] <- "ThreeAtmp"
    Basketball_Data_Filtered <- Basketball_Data_Filtered %>%
      mutate(ThreeAtemptPec = ThreeAtmp/FGA)

    p <- ggplot(data = Basketball_Data_Filtered,
                aes(x = as.numeric(season), y = ThreeAtemptPec,
                    text = paste('Year: ', season, '\n',
                                 '3PA / FGA: ', percent(ThreeAtmp/FGA, .01)))) +
      geom_point(color='salmon4') +
      labs(title = "3PA/FGA vs Season", x = "Season",
           y = "3PA / FGA")
    ggplotly(p, tooltip = "text")
    ```

## Predicting Future 3 Point Attempt Percentage

-   We then decided to predict what would the future 3 point attempt percentage based on the increase in years

-   We did this by using linear regression

-   We predicted for 10 years in the future (2032) and for 20 years in the future (2042)

    ```{r, echo=TRUE}
    colnames(results_final)[9] <- "ThreeAtmp"
    results_final <- results_final %>%
      mutate(ThreeAtemptPec = ThreeAtmp/FGA)

    lm_three_atempts <- lm(ThreeAtemptPec ~ season,
    data = results_final)

    summary(lm_three_atempts)
    ```

```{r, echo=TRUE}
#For 2032
Prediction_1 <- data.frame("season" = 2032)
predict(lm_three_atempts, newdata = Prediction_1)
```

```{r, echo=TRUE}
#For 2032
Prediction_2 <- data.frame("season" = 2042)
predict(lm_three_atempts, newdata = Prediction_2)
```

## Conclusion

-   Stephen Curry had a major impact on basketball 3 point shooting and the way the game is played today
-   We can also see from the descriptive statistics how different of a 3 point shooter he is compared to other pro basketball players
